---
title: "Assemble citation set for round 1 of classifier"
author: "O'Hara"
format: 
  html:
    code-fold: true
    embed-resources: true
execute:
  echo: true
  warning: false
  message: false
editor: source
---

This script gathers the citations and appends screening decisions to create a set of citations ready for the first round of the classifier.  

```{r setup}
library(tidyverse)
library(here)
library(synthesisr)
```


## Load all citations

Load the training set for the first pass of the classifier, and the remaining unscreened citations from Scopus and the Web of Science.

```{r}
training_df <- read_csv(here('int/classifier_training_set.csv'))

scopus_df <- read_refs(here('_data/3_refs_clean/scopus_clean_240204.ris')) %>%
  rename(key = notes)
wos_df <- read_refs(here('_data/3_refs_clean/wos_clean_240126.ris')) %>%
  rename(key = notes)
```

Bind the two datasets, and filter out any instances that appear in the `training_df`, using the key column.

```{r}
refs_df <- bind_rows(scopus_df, wos_df) %>%
  anti_join(training_df, by = 'key') %>%
  select(title, abstract, key)

write_csv(refs_df, here('int/classifier_round1_set.csv'))

### status 127 - can't find function `zip` on Windows - install RTools (outside of R)
zip(zipfile = here('int/classifier_round1_set.zip'), files = here('int/classifier_round1_set.csv'))
```

