---
title: "Implement LDA topic analysis"
author: "O'Hara"
format: 
  html:
    code-fold: true
    code-summary: "Show me the code"
execute:
  echo: true
  warning: false
  message: false
editor: source
---

```{r setup}
library(tidyverse)
library(here)
library(tidytext)
library(topicmodels)

source(here('common_fxns.R'))
```

Here we will implement a Latent Dirichlet Allocation analysis to help tease out topics from the corpus of articles gleaned from the Web of Science search.  For more on LDA:

> Blei, D. M. (2012). Probabilistic topic models. Communications of the ACM, 55(4), 77–84. https://doi.org/10.1145/2133806.2133826

For an example of topic modeling on which this script is based, see: https://www.tidytextmining.com/topicmodeling

## Load search results as document term matrix

```{r load bibtex}
abstr_df <- load_articles(pattern = 'wos_') %>%
  filter(topic == 'abstract')

abstr_words_df <- abstr_df %>%
  unnest_tokens(input = text, output = word, token = 'words') %>%
  anti_join(stop_words, by = 'word') %>%
  group_by(title, word) %>%
  summarize(n = n()) 

abstr_words_dtm <- abstr_words_df %>%
  cast_dtm(title, word, n)
```

## Perform LDA algorithm

How many latent topics to include?  How to identify an appropriate number of topics?  Examine processing time (work mac) and how it varies with $k$ and $n$.  

### Processing time as function of $n$ (number of articles)

As $n$ increases (more articles), the number of terms in the DTM will increase as well, so if LDA processing time is linear with matrix size, then it will be marginally increasing with $n$.

```{r how does term count change with n}
sample_abstr <- function(n, df = abstr_words_df) {
  titles_vec <- df$title %>% unique()
  titles_sample <- sample(titles_vec, size = n, replace = FALSE)
  out_df <- df %>%
    filter(title %in% titles_sample)  
  return(out_df)
}
words_per_sample <- function(n, df = abstr_words_df) {
  ### take the df = abstr_words_df, filter it for n articles, return word count
  term_count <- sample_abstr(n = n, df = df) %>%
    .$word %>% n_distinct()
  return(term_count)
}

n_vec <- 2^(0:11)
words_vs_n_df <- data.frame(n = n_vec,
                            term_count = sapply(n_vec, words_per_sample))
lm(term_count ~ sqrt(n), data = words_vs_n_df)

ggplot(words_vs_n_df, aes(x = n, y = term_count)) +
  geom_line() +
  geom_function(fun = function(x) -729 + 499 * sqrt(x), color = 'red', linetype = 'dashed')
```

Word count goes approximately as the square root of $n$, so doubling $n$ increases word count by a factor of ~1.4 (on top of doubling the number of articles).

```{r}
# abstr_sample_dtm <- sample_abstr(n = 3000) %>%
#   cast_dtm(title, word, n)
# system.time({
#   abstr_lda <- topicmodels::LDA(abstr_sample_dtm, k = 4, control = list(alpha = .1, seed = 4321))
# })
### on for k = 4: n = 200, 1.5 s; n = 400, 3.98 s; n = 600, 5.5 s; n = 1000, 15.9 s; n = 2000, 31.39 s; n = 3000, 39.0 s

p_df <- data.frame(n = c(200,  400, 600, 1000,  2000, 3000),
                   t = c(1.5, 3.98, 5.5, 15.9, 31.39, 39.0))
lm(t ~ n, p_df)

ggplot(p_df, aes(x = n, y = t)) + 
  geom_line() +
  geom_abline(intercept = -.845, slope = .014214, color = 'red', linetype = 'dashed') +
  labs(x = '# articles', y = 'processing time')
```

Processing time seems approximately linear with number of articles...

### Processing time as function of $k$ (number of topics)

```{r}
# abstr_sample_dtm <- sample_abstr(n = 1000) %>%
#   cast_dtm(title, word, n)
# system.time({
#   abstr_lda <- topicmodels::LDA(abstr_sample_dtm, k = 15, control = list(alpha = .1, seed = 4321))
# })
### on sample of 1000 articles: k = 2, 4.76 s; k = 4, 13.11s; k = 6, 22.92 s; k = 8, 35.17; k = 10, 40.33; k = 12, 47.98; k = 15, 56.26

p_df <- data.frame(k = c(   2,     4,     6,     8,    10,    12,    15),
                   t = c(4.76, 13.11, 22.92, 35.17, 40.33, 47.98, 56.26))

lm(t ~ k, p_df)

ggplot(p_df, aes(x = k, y = t)) + 
  geom_line() +
  geom_abline(intercept = -1.63, slope = 4.0691, color = 'red', linetype = 'dashed') +
  labs(x = '# topics', y = 'processing time')
```

Processing time seems approximately linear with number of topics... maybe tapering off toward the end there?

