---
title: "Systematic map: Benchmark test Web of Science"
author: "O'Hara"
format: 
  html:
    code-fold: true
    embed-resources: true
execute:
  echo: true
  warning: false
  message: false
editor: source
---

```{r}
library(tidyverse)
library(here)
library(tidytext)
library(bib2df)

source(here('common_fxns.R'))
```

## Check benchmarks against Web of Science search

```{r load bibtex}
fs <- list.files(here('_data/bibtex_clean'), pattern = 'wos_', full.names = TRUE)
all_fields_df <- lapply(fs, bib2df) %>% bind_rows()
```

```{r}
polish_records <- function(df) {
  df %>% 
    janitor::clean_names() %>%
    select(author, title, doi) %>%
    rowwise() %>%
    mutate(author = author[1] %>% str_remove(',.*')) %>%
    ungroup() %>%
    mutate(across(where(is.character), clean_text)) %>%
    ### remove foreign language translations of titles
    mutate(title = str_remove(title, '(; )?\\[.+\\] ?'))
}
```


```{r compare to bench}

result_df  <- all_fields_df %>%
  polish_records()

bench_df <- bib2df(here('_data/bibtex_clean/zot_benchmark_a.bib')) %>%
  polish_records() %>%
  mutate(title = str_remove_all(title, '\\{\\{|\\}\\}'))

doi_match <- bench_df %>%
  filter(!is.na(doi)) %>%
  filter(doi %in% result_df$doi)

info_match <- bench_df %>%
  select(author, title) %>%
  inner_join(result_df) %>%
  distinct()

match_df <- bind_rows(doi_match, info_match) %>%
  select(-doi) %>%
  distinct() %>%
  inner_join(result_df) %>%
  distinct()

pct_match <- nrow(match_df) / nrow(bench_df) * 100

missed <- bench_df %>% 
  select(author, title) %>%
  anti_join(match_df)

knitr::kable(missed %>% select(author, title))
```

## Check missed sources

Examining missed sources (Nov 20):

* Not in WoS database (white paper/grey lit?): Cooke, Bernknopf (whales), Kroetz, Mullan, Lombardo

## Hit rate

```{r}
bench_checked_df <- bench_df %>%
  filter(!author %in% c('cooke', 'kroetz', 'mullan', 'lombardo')) %>%
  filter(!(author == 'bernknopf' & str_detect(title, 'blue whale')))

hit_rate <- round(nrow(match_df) / nrow(bench_checked_df) * 100)

write_csv(match_df, here('_data/benchmark_matched.csv'))
```

There are `r nrow(bench_checked_df)` articles in the benchmark set that are known to be available in Web of Science.  The current search terms successfully identify these benchmark articles `r hit_rate`% of the time.


