---
title: "Examine results for round 2 of classifier"
author: "O'Hara"
format: 
  html:
    code-fold: true
    embed-resources: true
execute:
  echo: true
  warning: false
  message: false
editor: source
---

This script reads in the classifier model's results from round 2 and saves out a `.ris` of sampled citations for import into Colandr for citation screening, and then from these to fulltext screening.

This `.ris` will include all the predicted "includes".

```{r setup}
library(tidyverse)
library(here)
library(synthesisr)
```

## Load prediction results

```{r}
pred_df <- read_csv(here('_data/5d_classifier_round2', 
                         'predicted_classifier_round2_set.csv')) 
set.seed(42)
includes_df <- pred_df %>%
  filter(classification_prediction == 1) %>%
  rename(notes = key)

# sum(pred_df$classification_prediction) / nrow(pred_df) ### 16.5% hit rate far higher than likely
```

## Attach doi if available

Read in the full citations, grab the doi, and attach to the sampled citations. Shuffle the citations out to avoid unintentional bias

```{r}
scopus_df <- read_refs(here('_data/3_refs_clean/scopus_clean_240204.ris'))
wos_df <- read_refs(here('_data/3_refs_clean/wos_clean_240126.ris'))
refs_df <- bind_rows(scopus_df, wos_df) %>%
  select(notes, doi)

includes_out_df <- includes_df %>%
  mutate(source_type = 'article') %>%
  inner_join(refs_df, by = 'notes') %>%
  select(source_type, title, abstract, doi, notes) %>%
  slice_sample(prop = 1) %>%
  as.data.frame()

write_refs(includes_out_df, format = 'ris', 
           file = here('_data/5d_classifier_round2',
                       'predicted_classifier_round2_includes.ris'))
```
