---
title: "Pull all citations from round 3 of classifier"
author: "O'Hara"
format: 
  html:
    code-fold: true
    embed-resources: true
execute:
  echo: true
  warning: false
  message: false
editor: source
---

This script reads in the classifier model's results from round 3 and saves out a `.ris` of sampled citations for import into Colandr for citation screening, and then from these to fulltext screening.

```{r setup}
library(tidyverse)
library(here)
library(synthesisr)
```

## Load prediction results

```{r}
pred_df <- read_csv(here('_data/classifier_results', 
                         'predicted_classifier_round3_set.csv')) %>%
  rename(pred_round3 = 3, pred_round2 = 4)

table(pred_df %>% select(pred_round3, pred_round2))

incl_df <- pred_df %>% filter(pred_round3 == 1)

# sum(pred_df$pred_round3) / nrow(pred_df) ### 5.7%
```

Because all citations flagged as predicted includes for round 3 were also flagged in round 2, all round 3 citations are already in Colandr - no need for uploading additional references.
