---
title: "Network plot of authorship"
author: "O'Hara"
format: 
  html:
    code-fold: true
    embed-resources: true
execute:
  echo: true
  warning: false
  message: false
editor: source
---

```{r setup}
library(tidyverse)
library(here)
library(synthesisr)
library(igraph)
# library(ggnet2)

```

# Summary

Read in the includes data, including full author lists, and create a network plot of authorship.  This script follows tips from:

> Ognyanova, K. (2024) Network visualization with R. Retrieved from www.kateto.net/network-visualization.



# Methods

## Read and clean author fields for all "includes"

Keep only title and author fields.  The `key` field is not complete for early fulltext screening, but title should be unique.  Author format will be last name, first initial except where first initial is not unique (e.g., Rita and Roland Roberts), in which case first and middle initials will be used.

```{r}
clean_text <- function(t) {
  t %>%
    str_replace_all('\\\\~n', 'ñ') %>%   ### replace "\~n" with ñ
    str_replace_all("\\\\'a", 'á') %>%   ### replace "\~n" with ñ
    str_remove('\\\\relax') %>%          ### remove "\relax" from some author names
    str_remove_all('\\{|\\}') %>%        ### remove curly braces
    str_squish()
}

refs_raw <- read_refs(here('_data/screened_fulltext/convei_includes.bib'))

refs_clean <- refs_raw %>%
  select(title, author) %>%
  mutate(author = str_split(author, ' and ')) %>% 
  unnest(author) %>%
  ### establish first author status
  group_by(title) %>%
  mutate(lead_author = author == first(author)) %>%
  ungroup() %>%
  ### clean title and author names
  mutate(title_clean = clean_text(title)) %>%
  mutate(author = clean_text(author)) %>%
  ### separate name into last, first format
  separate(author, into = c('last', 'first'), sep = ',') %>%
  mutate(first = str_squish(first)) %>%
  ### fix first names to include middle init where available, for consistency
  mutate(first = case_when(last == 'Bernknopf' ~ 'Richard L',
                           last == 'Bouma' ~ 'Jetske A',
                           last == 'Dekker' ~ 'Arnold G',
                           last == 'Kuik' ~ 'Onno J',
                           last == 'Macauley' ~ 'Molly K',
                           last == 'Brookshire' ~ 'David S',
                           last == 'Podestá' ~ 'Guillermo P',
                           last == 'Raunikar' ~ 'R P',
                           last == 'Richardson' ~ 'Leslie A',
                           last == 'Seelan' ~ 'SK',
                           last == 'Solow' ~ 'Andrew R',
                           last == 'Weiher' ~ 'Rodney F',
                           last == 'Wilson' & first == 'J.' ~ 'James W',
                           last == 'Koontz' & first == 'Steve' ~ 'Stephen R',
                           last == 'Mishra' & first == 'Shruti' ~ 'Shruti K',
                           last == 'McCarl' & first == 'Bruce' ~ 'Bruce A',
                           last == "O'Brien" & first == 'James' ~ 'James J',
                           last == 'Miller' & first == 'Holly' ~ 'Holly M',
                           str_detect(title, 'Sydney 2000') & last == 'Roberts' ~ 'Rita D',
                           TRUE ~ first)) %>%
  mutate(first = str_remove_all(first, '[^A-Z]')) %>%
  mutate(first_letter = str_sub(first, 1, 1)) %>%
  ### if initial of first/middle name is unique, use only first initial
  group_by(last, first_letter) %>%
  mutate(n = n_distinct(first)) %>%
  ungroup() %>%
  mutate(first = ifelse(n == 1, first_letter, first)) %>%
  ### recombine first and last names into author id
  mutate(author = str_c(last, ', ', first)) %>%
  select(title = title_clean, author, lead_author)

### To simplify the plot, omit any papers where no authors have more than X papers,
### i.e., at least one author has multiple papers
refs_multi <- refs_clean %>%
  group_by(author) %>%
  mutate(n_papers = n_distinct(title)) %>%
  group_by(title) %>%
  filter(any(n_papers > 2)) %>%
  ungroup()
  
```



## Set up nodes and edges

Nodes are authors; edges are co-authorship relationships.  The `lead` field indicates whether the author is a lead author on at least one publication.

Annoyingly the ID columns are position-dependent not name dependent!

```{r}
nodes <- refs_multi %>%
  group_by(author) %>%
  summarize(lead = any(lead_author),
            n_papers = n()) %>%
  select(author, lead, n_papers) %>%
  mutate(label = ifelse(n_papers > 2, author, NA))

edges <- refs_multi %>%
  left_join(refs_clean, by = 'title') %>%
  ### discard same-author edges and reversed matches (e.g., A -> B and B -> A)
  filter(author.x > author.y) %>%
  select(from = author.x, to = author.y) %>%
  group_by(from, to) %>%
  summarize(weight = n())

```

Set up plot:

* Node size based on number of papers (not number of connections)
* Node color based on lead author status

```{r}
net <- graph_from_data_frame(d = edges, vertices = nodes, directed = FALSE) %>%
  simplify(remove.multiple = F, remove.loops = T)

### Set up vertex attributes
V(net)$size <- nodes$n_papers * 2
V(net)$color <- ifelse(nodes$lead, 'coral', 'gray80')
V(net)$label <- nodes$label

### set up edge attributes
E(net)$width <- E(net)$weight^2 / 2
# edge_pal <- c('grey95', 'grey90', 'grey85', 'grey80')
# E(net)$color <- edge_pal[E(net)$weight]
E(net)$color <- 'grey80'

plot(net, edge.curved = 0.1,
     vertex.label.family = 'sans', vertex.label.color = 'black', vertex.label.cex = 0.6,
     # vertex.label.dist = 1.5, vertex.label.degree = 0,
     vertex.frame.color = NA,
     layout = layout_nicely)
# legend(x="bottomleft", c("Lead author", "none"), pch=21,
#        col=NA, pt.bg=c('coral', 'gray80'), pt.cex=2, cex=.8, bty="n", ncol=1)

```

