---
title: "Analysis: final include set: location of studies"
author: "O'Hara"
format: 
  html:
    code-fold: true
    embed-resources: true
execute:
  echo: true
  warning: false
  message: false
editor: source
---

```{r setup}
library(tidyverse)
library(here)
library(sf)
```

# Summary

Read in all the final "includes" and summarize by:

* societal benefit area
* societal benefit type
* study location

For papers where two methods, societal benefit areas, or societal benefit types were coded, we will count each instance separately.  This may be a point of discussion - perhaps they are weighted to add up to a single instance?

# Methods

Read in coded data of included papers.

```{r}
includes_df <- readxl::read_excel(here('_data/screened_fulltext/ft_consolidated_coding.xlsx'))
```

### Summarize by study location

```{r}
loc_df <- includes_df %>%
  select(key, location = 'study location') %>%
  pivot_longer(-key, names_to = 'tmp', values_to = 'loc') %>%
  filter(!is.na(loc)) %>%
  select(-tmp) %>%
  distinct() %>%
  mutate(loc = str_split(loc, '[^A-z ]| and ')) %>%
  unnest(loc) %>%
  mutate(loc = str_squish(loc)) %>%
  filter(!loc %in% c('', 'case')) %>%
  group_by(loc) %>%
  mutate(loc = ifelse(n() == 1, 'Other', loc)) %>%
  ungroup()

loc_summary <- loc_df %>%
  summarize_column('loc') %>%
  mutate(loc = fct_relevel(loc, 'Other', after = 0))

gen_plot(loc_summary, 'loc', 'Study location')
```

```{r}
georgns <- read_csv('https://raw.githubusercontent.com/lukes/ISO-3166-Countries-with-Regional-Codes/refs/heads/master/all/all.csv') %>%
  janitor::clean_names() %>%
  mutate(name = str_replace(name, 'United States of America', 'United States'),
         name = str_replace(name, 'United Kingdom of Great Britain and Northern Ireland', 'United Kingdom'),
         name = str_replace(name, 'Syrian Arab Republic', 'Syria'),
         name = str_replace(name, 'North Macedonia', 'Macedonia'),
         name = ifelse(alpha_2 == 'KR', 'South Korea', name),
         name = str_remove(name, ',.+'))


loc2_df <- includes_df %>%
  select(key, location = 'study location') %>%
  pivot_longer(-key, names_to = 'tmp', values_to = 'loc') %>%
  filter(!is.na(loc)) %>%
  select(-tmp) %>%
  distinct() %>%
  mutate(loc = str_split(loc, '[^A-z ]')) %>%
  unnest(loc) %>%
  filter(!loc %in% c('', 'case')) %>%
  mutate(loc = str_squish(loc)) %>%
  left_join(georgns, by = c('loc' = 'name')) %>%
  mutate(loc2 = case_when(loc == 'United States' ~ 'United States',
                          loc == 'Taiwan' ~ 'Eastern Asia',
                          is.na(sub_region) ~ loc,
                          TRUE ~ sub_region))

loc2_summary <- loc2_df %>%
  select(key, loc2) %>%
  distinct() %>%
  group_by(loc2) %>%
  summarize(n = n())
  

```



