---
title: "Sample predicted excludes for Colandr"
author: "O'Hara"
format: 
  html:
    code-fold: true
    embed-resources: true
execute:
  echo: true
  warning: false
  message: false
editor: source
---

This script reads in the classifier model's results from round 2 and saves out a `.ris` of sampled citations for import into Colandr for citation screening, and then from these to fulltext screening.  

This `.ris` will 1000 predicted "excludes" from the 12532 remaining citations.

```{r setup}
library(tidyverse)
library(here)
library(synthesisr)
```

## Load prediction results

```{r}
pred_df <- read_csv(here('_data/classifier_results', 
                         'predicted_classifier_round2_set.csv')) 

excl_df <- pred_df %>% filter(classification_prediction == 0)

# sum(pred_df$classification_prediction) / nrow(pred_df) ### 17.1% hit rate far higher than likely
```

## Attach doi if available

Read in the full citations, grab the doi, and attach to the sampled citations.  Shuffle the citations out to avoid unintentional bias

```{r}
scopus_df <- read_refs(here('_data/1c_refs_clean/scopus_clean_240204.ris'))
wos_df <- read_refs(here('_data/1c_refs_clean/wos_clean_240126.ris'))
refs_df <- bind_rows(scopus_df, wos_df) %>%
  select(source_type, title, author, year, abstract, journal, notes, doi)

set.seed(42)
colandr_out_df <- excl_df %>%
  rename(notes = key) %>% ### for consistency with RIS
  inner_join(refs_df, by = c('title', 'notes')) %>%
  select(source_type, title, abstract, author, year, doi, notes) %>%
  as.data.frame() %>%
  slice_sample(n = 1000)

write_refs(colandr_out_df, format = 'ris', 
           file = here('_data/6_screen_classifier_excludes',
                       'ris_to_colandr_excl_sample.ris'))
```


