---
title: "Surveying the landscape of approaches to assessing societal benefits of Earth science information: a systematic map"
author: Casey C. O'Hara (https://orcid.org/0000-0003-2968-7005)^1\*^,
        Samantha Cheng^2^,
        Mabel Baez-Schon^2^,
        Rebecca Chaplin-Kramer^2^,
        Alejandra Echeverri (https://orcid.org/0000-0002-8821-0509)^3^,
        Gillian Galford (https://orcid.org/0000-0003-2192-7385)^4,5^,
        Rachelle K. Gould (https://orcid.org/0000-0002-6307-8783)^4,5^,
        Cristina Mancilla^1^,
        Maura Muldoon (https://orcid.org/0009-0007-8151-3164)^4,5^,
        Gerald Singh (https://orcid.org/0000-0003-4333-1988)^6^,
        Priscilla Baltezar^7^,
        Yusuke Kuwayama^8^,
        Stephen Polasky (https://orcid.org/0000-0003-4934-2434)^9^,
        Amanda D. Rodewald (https://orcid.org/0000-0002-6719-6306)^10^,
        Elizabeth Tennant (https://orcid.org/0000-0003-3858-9693)^11^,
        Jiaying Zhao (https://orcid.org/0000-0003-0902-4343)^12^,
        Benjamin S. Halpern^1,13^
csl: refs/pnas.csl
execute:
  echo: false
  warning: false
  message: false
format:
  docx: 
    reference-doc: templates/msword_pnas_template.docx
filters:
  - refs/docx-landscape.lua
bibliography: refs/sp1_ms_refs.bib
---

``` {r setup}
#| eval: false
```

<!-- Include, in this order, department/laboratory/section/division, institution, city, state with ZIP code (for US institutions) or country with postal code (for non-US institutions). -->
<p>

**Short Title:** TBD

### Affiliations

1. National Center for Ecological Analysis and Synthesis, Santa Barbara CA 93101, USA
2. Global Science, World Wildlife Fund, Washington, DC, 10024, USA
3. Department of Environmental Science, Policy and Management. University of California, Berkeley, Berkeley CA, 94709
4. Rubenstein School of Environment and Natural Resources, University of Vermont, Burlington VT 05401, USA
5. Gund Institute for Environment, University of Vermont, Burlington VT 05401, USA
6. School of Environmental Studies, University of Victoria, Victoria BC V8P 5C2, Canada
7. Massachusetts Institute of Technology Space Enabled Research Group, Cambridge MA 02139, USA
8. School of Public Policy, University of Maryland, Baltimore County, Baltimore MD 21250, USA
9. Department of Applied Economics, University of Minnesota, St. Paul MN 55108, USA
10. Cornell Lab of Ornithology and Department of Natural Resources and the Environment, Cornell University, Ithaca NY 14850, USA
11. Charles H. Dyson School, Cornell University, Ithaca NY 14853, USA
12. Department of Psychology, Institute for Resources, Environment and Sustainability, University of British Columbia, Vancouver BC V6Z 1T4, Canada
13. Bren School of Environmental Science & Management, UC Santa Barbara, Santa Barbara CA 93106, USA



__^\*^Correspondence to:__  Casey O'Hara, ohara@nceas.ucsb.edu

<!-- Include the name, complete address, phone number, and email address of the author(s) to whom correspondence and proofs should be sent. -->

<!-- __Competing Interest Statement:__ The authors have no competing interests.   -->

<!-- __Classification:__ Biological Sciences: Sustainability Science  -->

__Keywords:__ TBD


## This file includes:

* Main Text
* Figures 1-6
* References 1-63

\pagebreak

# Abstract

``` {r abstract}
#| child: '0_abstract.qmd'
```

# Main Text

## Introduction

``` {r intro}
#| child: '1_intro.qmd'
```


## Results

``` {r results}
#| child: '2_results.qmd'
```

## Discussion

``` {r discussion}
#| child: '3_discussion.qmd'
```

## Conclusion

``` {r conclusion}
#| child: '4_conclusion.qmd'
```

## Methods

``` {r results}
#| child: '5_methods.qmd'
```

## Author Contributions

Conceptualization, TBD; Methodology, TBD; Software, CCO; Formal Analysis, CCO; Data Curation, CCO; Writing - Original Draft, CCO; Writing - Review & Editing, CCO, TBD; Visualization, CCO; Supervision, BSH, BCK; Funding Acquisition, BSH, BCK.

## Acknowledgments

TBD We thank the National Center for Ecological Analysis and Synthesis (NCEAS) for computational support.  <!--__Funding__: We gratefully acknowledge financial support from NCEAS, the National Science Foundation, the National Philanthropic Trust, and a fellowship from the Bren School of Environmental Science and Management.-->



\pagebreak
# References

<div id="refs"></div>



<!-- \pagebreak -->
<!-- # Supporting information -->

``` {r SI}
#| eval: true
#| child: '_SI.qmd'
```
